---
title: "GC Workflow (All Runs): Peak Heights → PPM"
author: "Gavin McNicol"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 2
runtime: shiny
params:
  gc_raw_root: "flux-processing/data/fluxes/gc-raw"
  tidy_root: "flux-processing/data/tidy"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(readr)
library(janitor)
library(lubridate)
library(stringr)
library(DT)
library(shiny)

# source functions
# source("flux-processing/R/gc_batch_process.R")
source("flux-processing/R/heights_to_ppm.R")
source("flux-processing/R/ppm_to_amount.R")
source("flux-processing/R/flag_and_drop_bad_jumps_iterative.R")
source("flux-processing/R/flag_and_drop_bad_jumps.R")


# source("flux-processing/run-step3-6.R")

gc_raw_root <- params$gc_raw_root
tidy_root   <- params$tidy_root

```

# STEP ONE - GET PEAK HEIGHTS AND MATCH SAMPLES/STDs USING SAMPLE_LOG.csv

For all dates (folders)...

```{r all-files}
log_msg("GC raw root:", gc_raw_root)
log_msg("Tidy root:", tidy_root)

# Find all directories containing a sample_log.csv
run_dirs <- list.dirs(gc_raw_root, recursive = TRUE, full.names = TRUE)
run_dirs <- run_dirs[file.exists(file.path(run_dirs, "sample_log.csv"))]

log_msg("Found", length(run_dirs), "run directories with sample_log.csv")

# Process each run dir -> writes *_gc_peak_heights.csv (and your *_gc_standards.csv if enabled)
walk(run_dirs, ~process_gc_run(.x, tidy_root = tidy_root))
```

Or for a single, selected date (folder)...

```{r select-date-folder}
# # Pick a date to run
# selected_date <- as.Date("2023-07-18")
# 
# run_dirs <- list.dirs(gc_raw_root, recursive = TRUE, full.names = TRUE)
# run_dirs <- run_dirs[file.exists(file.path(run_dirs, "sample_log.csv"))]
# 
# run_dirs <- run_dirs[
#   !is.na(lubridate::mdy(stringr::str_replace_all(basename(run_dirs), "\\.", "/"))) &
#   lubridate::mdy(stringr::str_replace_all(basename(run_dirs), "\\.", "/")) == selected_date
# ]
# 
# log_msg("Selected date:", as.character(selected_date))
# log_msg("Runs to process:", length(run_dirs))
# 
# walk(run_dirs, ~process_gc_run(.x, tidy_root = tidy_root))
```

Check sampling date .csv for bad dates.

```{r check-dates}
peak_summary <- list.files(
  "flux-processing/data/tidy",
  pattern = "_gc_peak_heights\\.csv$",
  recursive = TRUE,
  full.names = TRUE
) %>%
  tibble(file = .) %>%
  mutate(
    run_label = basename(dirname(file))
  ) %>%
  mutate(
    data = map(file, ~ suppressWarnings(readr::read_csv(.x, show_col_types = FALSE)))
  ) %>%
  unnest(data) %>%
  group_by(run_label) %>%
  summarize(
    n_rows = n(), # number of rows in gc run(s) for that date
    n_na_meta = sum(is.na(site) | is.na(gc_run_date)), # number of NAs in metadata columns  (for same date)
    n_na_heights = sum(is.na(co2_height) | is.na(ch4_height) | is.na(n2o_height)), # number of na peak heights (for same date)
    has_stds = any(str_detect(as.character(chamber), regex("std", ignore_case = TRUE))), # does gc run(s) include standards
    .groups = "drop"
  ) %>%
  arrange(run_label)

peak_summary

peak_summary %>%
  filter(n_na_meta == n_rows) %>%
  pull(run_label)
```

Check specific date tables.

```{r}
run <- "220722"

ph <- read_csv(glue::glue("flux-processing/data/tidy/{run}/{run}_gc_peak_heights.csv"))
sl <- read_csv(glue::glue("flux-processing/data/fluxes/gc-raw/2022/7.22.22/sample_log.csv")) %>%
  janitor::clean_names()

setdiff(unique(ph$filename), unique(sl$filename))
setdiff(unique(sl$filename), unique(ph$filename))
```

Secondary data check... output only contains dates with potential issues.

  * Main concern is if `all_` columns are both TRUE. 

```{r identify-bad-processing}
peak_files <- list.files(
  "flux-processing/data/tidy",
  pattern = "_gc_peak_heights\\.csv$",
  full.names = TRUE,
  recursive = TRUE
)

meta_cols_strict <- c("filename", "gc_run_date", "site", "timepoint")
meta_cols_nonstd <- c("date_collected", "ecosystem", "chamber_no")
height_cols      <- c("co2_height", "ch4_height", "n2o_height")

# helper: pull a set of columns, adding missing ones as NA
get_cols_or_na <- function(df, cols) {
  missing <- setdiff(cols, names(df))
  if (length(missing)) {
    df[missing] <- NA
  }
  df[, cols, drop = FALSE]
}

peak_qc <- tibble(file = peak_files) %>%
  mutate(
    dat = map(file, ~ readr::read_csv(.x, show_col_types = FALSE)),

    n_rows = map_int(dat, nrow),

    any_na_heights = map_lgl(dat, ~ {
      h <- get_cols_or_na(.x, height_cols)
      any(is.na(h))
    }),
    all_na_heights = map_lgl(dat, ~ {
      h <- get_cols_or_na(.x, height_cols)
      all(is.na(h))
    }),

    # strict meta: must exist + non-NA for everyone (i.e., flag if ANY NA anywhere)
    any_na_meta_strict = map_lgl(dat, ~ {
      m <- get_cols_or_na(.x, meta_cols_strict)
      any(is.na(m))
    }),
    # "all rows have all strict meta NA" (or columns missing => NA) => TRUE
    all_na_meta_strict = map_lgl(dat, ~ {
      m <- get_cols_or_na(.x, meta_cols_strict)
      all(rowSums(!is.na(m)) == 0)
    }),

    # non-STD meta: only enforce for non-standards
    any_na_meta_nonstd = map_lgl(dat, ~ {
      df <- .x

      # ensure site/chamber exist for STD detection
      if (!"site" %in% names(df)) df$site <- NA_character_
      if (!"chamber" %in% names(df)) df$chamber <- NA_character_

      is_std <- grepl("std", df$site, ignore.case = TRUE) |
                grepl("std", df$chamber, ignore.case = TRUE)

      x <- df[!is_std %in% TRUE, , drop = FALSE]  # keep FALSE/NA as "non-STD-ish"
      if (nrow(x) == 0) return(FALSE)

      m2 <- get_cols_or_na(x, meta_cols_nonstd)
      any(is.na(m2))
    })
  ) %>%
  select(-dat)

peak_qc %>%
  filter(
    any_na_heights | all_na_heights |
      any_na_meta_strict | all_na_meta_strict |
      any_na_meta_nonstd
  ) %>%
  arrange(desc(any_na_meta_nonstd), desc(any_na_meta_strict), desc(any_na_heights)) %>%
  print(n = Inf)
```

# STEP 2 - CONVERT HEIGHTS TO PPM and OUTPUT SAMPLE PPM and STD PPM

Note: chamber breakdown of unknown counts may be strange, but is addressed later.

```{r}
# tidy_root should match what you used in step 01
tidy_root <- "flux-processing/data/tidy"

peak_files <- list.files(
  tidy_root,
  pattern = "_gc_peak_heights\\.csv$",
  full.names = TRUE,
  recursive = TRUE
)

# helper: infer run_label from folder name (e.g. .../230802/230802_gc_peak_heights.csv)
get_run_label <- function(path) basename(dirname(path))

ppm_batch <- tibble(
  run_label = map_chr(peak_files, get_run_label),
  peaks_file = peak_files
) %>%
  arrange(run_label) %>%
  mutate(
    result = map(peaks_file, ~{
      tryCatch(
        heights_to_ppm(.x),  # uses your default output naming in same folder
        error = function(e) e
      )
    }),
    ok = map_lgl(result, ~!inherits(.x, "error")),
    err = map_chr(result, ~if (inherits(.x, "error")) conditionMessage(.x) else NA_character_)
  ) %>%
  select(-result)

# Show any failures
ppm_batch %>% filter(!ok) %>% print(n = Inf)

# Quick success count
ppm_batch %>% count(ok)
```

Review complete `_ghg_ppm.csv` file outputs. Length should match number of sampling dates.

```{r review-file}
# tidy_root should match what you used in step 01
tidy_root <- "flux-processing/data/tidy"

peak_files <- list.files(
  tidy_root,
  pattern = "_ghg_ppm\\.csv$",
  full.names = TRUE,
  recursive = TRUE
)

peak_files
```

# STEP THREE - FIVE - FIT LINEAR MODELS AND GET SIGNIFICANT SLOPES

```{r}
date <- "220902"
run_date_pipeline(run_label = date,
                  k_jump = 1,
                  k_out = 1)
```

# inspect all bad fits

```{r}
library(tidyverse)

tidy_root <- "flux-processing/data/tidy"

# ---- set the date you want (YYMMDD)
run_label_target <- date

# 1) Find QC files and keep only the specified date
qc_files_all <- list.files(
  tidy_root,
  pattern = "_flux_fits_qc\\.csv$",
  full.names = TRUE,
  recursive = TRUE
)

qc_files <- qc_files_all %>%
  keep(~ str_detect(.x, paste0("/", run_label_target, "/"))) %>%  # folder filter
  keep(~ str_detect(basename(.x), paste0("^", run_label_target))) # filename sanity check

if (length(qc_files) == 0) {
  stop("No QC files found for run_label_target = ", run_label_target,
       " under ", tidy_root)
}

# 2) Read + bind, then filter non-OK
fits_nonok <- tibble(qc_file = qc_files) %>%
  mutate(
    run_label = run_label_target,
    dat = map(qc_file, ~ readr::read_csv(.x, show_col_types = FALSE))
  ) %>%
  unnest(dat) %>%
  mutate(
    qc_flag = as.character(qc_flag),
    gas     = as.character(gas)
  ) %>%
  filter(!is.na(qc_flag), qc_flag != "OK") %>%
  select(run_label, qc_flag, everything())

# 3) Summaries
fits_nonok %>%
  count(gas, qc_flag, sort = TRUE) %>%
  print(n = Inf)

fits_nonok %>%
  count(run_label, gas, qc_flag, sort = TRUE) %>%
  print(n = 50)
```

# plot all fits

```{r}
library(tidyverse)

tidy_root <- "flux-processing/data/tidy"

# ---- set the date you want (YYMMDD)
run_label_target <- date

# Options
gases_to_plot <- c("CO2_umol", "CH4_nmol", "N2O_nmol")
save_plots <- TRUE

plot_dir <- file.path(tidy_root, "_qc_plots", run_label_target)
dir.create(plot_dir, showWarnings = FALSE, recursive = TRUE)

# 1) Find QC + amount files for ONLY this date
qc_files_all  <- list.files(tidy_root, pattern = "_flux_fits_qc\\.csv$", full.names = TRUE, recursive = TRUE)
amt_files_all <- list.files(tidy_root, pattern = "_ghg_amount\\.csv$",    full.names = TRUE, recursive = TRUE)

qc_files <- qc_files_all %>%
  keep(~ str_detect(.x, paste0("/", run_label_target, "/"))) %>%
  keep(~ str_detect(basename(.x), paste0("^", run_label_target)))

amt_files <- amt_files_all %>%
  keep(~ str_detect(.x, paste0("/", run_label_target, "/"))) %>%
  keep(~ str_detect(basename(.x), paste0("^", run_label_target)))

if (length(qc_files) == 0)  stop("No QC files found for ", run_label_target)
if (length(amt_files) == 0) stop("No amount files found for ", run_label_target)

read_csv0 <- function(f) readr::read_csv(f, show_col_types = FALSE)

# 2) Read QC fits (date-restricted)
fits_qc <- purrr::map_dfr(qc_files, read_csv0) %>%
  mutate(
    date = as.character(date),
    ecosystem_name = as.character(ecosystem_name),
    ecosystem_block = as.integer(ecosystem_block),
    rep = as.integer(rep),
    gas = as.character(gas),
    qc_flag = as.character(qc_flag)
  )

# 3) Read amount rows (date-restricted), pivot longer to (gas, amount)
amount_long <- purrr::map_dfr(amt_files, read_csv0) %>%
  select(date, ecosystem_name, ecosystem_block, rep, time_point, min_elapsed,
         co2_umol, ch4_nmol, n2o_nmol) %>%
  mutate(
    date = as.character(date),
    ecosystem_name = as.character(ecosystem_name),
    ecosystem_block = as.integer(ecosystem_block),
    rep = as.integer(rep),
    time_point = as.integer(time_point),
    min_elapsed = as.numeric(min_elapsed)
  ) %>%
  pivot_longer(
    cols = c(co2_umol, ch4_nmol, n2o_nmol),
    names_to = "gas_col",
    values_to = "amount"
  ) %>%
  mutate(
    amount = as.numeric(amount),
    gas = recode(gas_col,
                 co2_umol = "CO2_umol",
                 ch4_nmol = "CH4_nmol",
                 n2o_nmol = "N2O_nmol")
  ) %>%
  select(-gas_col)

# 4) Join amount to QC for ALL chamber-series (OK + non-OK)
plot_df_all <- amount_long %>%
  left_join(
    fits_qc,
    by = c("date", "ecosystem_name", "ecosystem_block", "rep", "gas")
  ) %>%
  mutate(
    qc_flag = replace_na(qc_flag, "NO_QC_ROW"),
    chamber_id = paste0(ecosystem_name, "-", ecosystem_block,
                        " Rep ", rep, " (", date, ")"),
    qc_label = paste0(qc_flag,
                      " | R2=", ifelse(is.na(r.squared), "NA", signif(r.squared, 3)),
                      " | p=",  ifelse(is.na(p_value),   "NA", signif(p_value, 3)),
                      " | slope=", ifelse(is.na(slope), "NA", signif(slope, 3)))
  )

# 5) Plot helper
make_gas_plot <- function(df, gas_name) {
  d <- df %>% filter(gas == gas_name)

  if (nrow(d) == 0) {
    message("No data found for gas: ", gas_name)
    return(NULL)
  }

  d <- d %>%
    mutate(
      qc_rank = case_when(
        qc_flag == "OK" ~ 2L,
        qc_flag == "NO_QC_ROW" ~ 3L,
        TRUE ~ 1L
      )
    ) %>%
    arrange(qc_rank, chamber_id) %>%
    mutate(chamber_id = factor(chamber_id, levels = unique(chamber_id)))

  ggplot(d, aes(x = min_elapsed, y = amount)) +
    geom_point(aes(shape = factor(time_point), color = qc_flag), size = 3, alpha = 0.9) +
    geom_smooth(method = "lm", se = FALSE, linewidth = 0.7, color = "black") +
    facet_wrap(~ chamber_id) +
    labs(
      title = paste0("All chamber series (QC-colored): ", gas_name, " | ", run_label_target),
      subtitle = "Black line = LM fit per chamber-series; points colored by qc_flag (OK vs non-OK).",
      x = "Minutes elapsed",
      y = gas_name,
      color = "QC flag",
      shape = "Timepoint"
    ) +
    theme_bw() +
    theme(
      strip.text = element_text(size = 10),
      axis.text = element_text(size = 10),
      legend.position = "bottom"
    )
}

# 6) Generate plots for each gas
plots <- purrr::map(gases_to_plot, ~ make_gas_plot(plot_df_all, .x))
names(plots) <- gases_to_plot

purrr::walk(plots, ~ if (!is.null(.x)) print(.x))

# Save each gas to its own PDF
if (isTRUE(save_plots)) {
  purrr::iwalk(plots, function(p, gas_name) {
    if (is.null(p)) return(invisible(NULL))
    out_pdf <- file.path(plot_dir, paste0("qc_all_chambers_", run_label_target, "_", gas_name, ".pdf"))
    ggsave(out_pdf, plot = p, width = 14, height = 10, units = "in")
    message("Wrote: ", out_pdf)
  })
}
```
 
```{r}
read_csv("flux-processing/data/tidy/220812/220812_flux_fits_flux.csv")
```

## DONE WITH CODE FOR GENERATING FLUXES, just need to do it now for all the dates, two issues: missing min_elapsed and temps for first two dates in 2022, and need to make sure date is correct in chamber_temps through all datse (noticed some duplicatoin)
 

