---
title: "GC Workflow (All Runs): Peak Heights → PPM"
author: "Gavin McNicol"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 2
runtime: shiny
params:
  gc_raw_root: "flux-processing/data/fluxes/gc-raw"
  tidy_root: "flux-processing/data/tidy"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(readr)
library(janitor)
library(lubridate)
library(stringr)
library(DT)
library(shiny)

# source functions
source("flux-processing/gc_batch_process.R")
source("flux-processing/gc_heights_to_ppm.R")

gc_raw_root <- params$gc_raw_root
tidy_root   <- params$tidy_root

```

# STEP ONE - GET PEAK HEIGHTS AND MATCH SAMPLES/STDs USING SAMPLE_LOG.csv

```{r all-files}
log_msg("GC raw root:", gc_raw_root)
log_msg("Tidy root:", tidy_root)

# Find all directories containing a sample_log.csv
run_dirs <- list.dirs(gc_raw_root, recursive = TRUE, full.names = TRUE)
run_dirs <- run_dirs[file.exists(file.path(run_dirs, "sample_log.csv"))]

log_msg("Found", length(run_dirs), "run directories with sample_log.csv")

# Process each run dir -> writes *_gc_peak_heights.csv (and your *_gc_standards.csv if enabled)
walk(run_dirs, ~process_gc_run(.x, tidy_root = tidy_root))
```

```{r select-date-folder}
# Pick a date to run
selected_date <- as.Date("2023-07-18")

run_dirs <- list.dirs(gc_raw_root, recursive = TRUE, full.names = TRUE)
run_dirs <- run_dirs[file.exists(file.path(run_dirs, "sample_log.csv"))]

run_dirs <- run_dirs[
  !is.na(lubridate::mdy(stringr::str_replace_all(basename(run_dirs), "\\.", "/"))) &
  lubridate::mdy(stringr::str_replace_all(basename(run_dirs), "\\.", "/")) == selected_date
]

log_msg("Selected date:", as.character(selected_date))
log_msg("Runs to process:", length(run_dirs))

walk(run_dirs, ~process_gc_run(.x, tidy_root = tidy_root))
```

Check sampling date .csv for bad dates.

```{r check-dates}
peak_summary <- list.files(
  "flux-processing/data/tidy",
  pattern = "_gc_peak_heights\\.csv$",
  recursive = TRUE,
  full.names = TRUE
) %>%
  tibble(file = .) %>%
  mutate(
    run_label = basename(dirname(file))
  ) %>%
  mutate(
    data = map(file, ~ suppressWarnings(readr::read_csv(.x, show_col_types = FALSE)))
  ) %>%
  unnest(data) %>%
  group_by(run_label) %>%
  summarize(
    n_rows = n(),
    n_na_meta = sum(is.na(site) | is.na(gc_run_date)),
    n_na_heights = sum(is.na(co2_height) | is.na(ch4_height) | is.na(n2o_height)),
    has_stds = any(str_detect(as.character(chamber), regex("std", ignore_case = TRUE))),
    .groups = "drop"
  ) %>%
  arrange(run_label)

peak_summary

peak_summary %>%
  filter(n_na_meta == n_rows) %>%
  pull(run_label)
```

Check specific date tables.

```{r}
run <- "230818"

ph <- read_csv(glue::glue("flux-processing/data/tidy/{run}/{run}_gc_peak_heights.csv"))
sl <- read_csv(glue::glue("flux-processing/data/fluxes/gc-raw/2023/8.18.23/sample_log.csv")) %>%
  janitor::clean_names()

setdiff(unique(ph$filename), unique(sl$filename))
setdiff(unique(sl$filename), unique(ph$filename))
```


```{r identify-bad-processing}
peak_files <- list.files(
  "flux-processing/data/tidy",
  pattern = "_gc_peak_heights\\.csv$",
  full.names = TRUE,
  recursive = TRUE
)

meta_cols_strict <- c("filename", "gc_run_date", "site", "timepoint")
meta_cols_nonstd <- c("date_collected", "ecosystem", "chamber_no")
height_cols      <- c("co2_height", "ch4_height", "n2o_height")

# helper: pull a set of columns, adding missing ones as NA
get_cols_or_na <- function(df, cols) {
  missing <- setdiff(cols, names(df))
  if (length(missing)) {
    df[missing] <- NA
  }
  df[, cols, drop = FALSE]
}

peak_qc <- tibble(file = peak_files) %>%
  mutate(
    dat = map(file, ~ readr::read_csv(.x, show_col_types = FALSE)),

    n_rows = map_int(dat, nrow),

    any_na_heights = map_lgl(dat, ~ {
      h <- get_cols_or_na(.x, height_cols)
      any(is.na(h))
    }),
    all_na_heights = map_lgl(dat, ~ {
      h <- get_cols_or_na(.x, height_cols)
      all(is.na(h))
    }),

    # strict meta: must exist + non-NA for everyone (i.e., flag if ANY NA anywhere)
    any_na_meta_strict = map_lgl(dat, ~ {
      m <- get_cols_or_na(.x, meta_cols_strict)
      any(is.na(m))
    }),
    # "all rows have all strict meta NA" (or columns missing => NA) => TRUE
    all_na_meta_strict = map_lgl(dat, ~ {
      m <- get_cols_or_na(.x, meta_cols_strict)
      all(rowSums(!is.na(m)) == 0)
    }),

    # non-STD meta: only enforce for non-standards
    any_na_meta_nonstd = map_lgl(dat, ~ {
      df <- .x

      # ensure site/chamber exist for STD detection
      if (!"site" %in% names(df)) df$site <- NA_character_
      if (!"chamber" %in% names(df)) df$chamber <- NA_character_

      is_std <- grepl("std", df$site, ignore.case = TRUE) |
                grepl("std", df$chamber, ignore.case = TRUE)

      x <- df[!is_std %in% TRUE, , drop = FALSE]  # keep FALSE/NA as "non-STD-ish"
      if (nrow(x) == 0) return(FALSE)

      m2 <- get_cols_or_na(x, meta_cols_nonstd)
      any(is.na(m2))
    })
  ) %>%
  select(-dat)

peak_qc %>%
  filter(
    any_na_heights | all_na_heights |
      any_na_meta_strict | all_na_meta_strict |
      any_na_meta_nonstd
  ) %>%
  arrange(desc(any_na_meta_nonstd), desc(any_na_meta_strict), desc(any_na_heights)) %>%
  print(n = Inf)
```

# STEP 2 - CONVERT HEIGHTS TO PPM and OUTPUT SAMPLE PPM and STD PPM

```{r}
# tidy_root should match what you used in step 01
tidy_root <- "flux-processing/data/tidy"

peak_files <- list.files(
  tidy_root,
  pattern = "_gc_peak_heights\\.csv$",
  full.names = TRUE,
  recursive = TRUE
)

# helper: infer run_label from folder name (e.g. .../230802/230802_gc_peak_heights.csv)
get_run_label <- function(path) basename(dirname(path))

ppm_batch <- tibble(
  run_label = map_chr(peak_files, get_run_label),
  peaks_file = peak_files
) %>%
  arrange(run_label) %>%
  mutate(
    result = map(peaks_file, ~{
      tryCatch(
        heights_to_ppm(.x),  # uses your default output naming in same folder
        error = function(e) e
      )
    }),
    ok = map_lgl(result, ~!inherits(.x, "error")),
    err = map_chr(result, ~if (inherits(.x, "error")) conditionMessage(.x) else NA_character_)
  ) %>%
  select(-result)

# Show any failures
ppm_batch %>% filter(!ok) %>% print(n = Inf)

# Quick success count
ppm_batch %>% count(ok)
```

```{r review-file}
# tidy_root should match what you used in step 01
tidy_root <- "flux-processing/data/tidy"

peak_files <- list.files(
  tidy_root,
  pattern = "_ghg_ppm\\.csv$",
  full.names = TRUE,
  recursive = TRUE
)

peak_files
```


## Overview

This RMarkdown runs the GC processing workflow in two steps for a single run date:
	1.	Extract peak heights from all .txt files in the GC run folder and write a tidy CSV:
	•	{tidy_root}/{run_label}/{run_label}_gc_peak_heights.csv
	2.	Convert peak heights to ppm (using standards) and write:
	•	{r  un_label}_std_heights.csv
	•	{run_label}_ghg_ppm.csv (only if non-STD samples exist)

Example run folder: flux-processing/data/fluxes/gc-raw/2022/7.9.22

## 1 - Extract peak heights for 7.9.22

```{r run-step-1}
run_dir   <- "flux-processing/data/fluxes/gc-raw/2022/7.9.22"
tidy_root <- "flux-processing/data/tidy"

res_peaks <- process_gc_run(run_dir, tidy_root = tidy_root)
```

Inspect written peak heights

```{r inspect-peaks}
run_label  <- "220709"
peaks_file <- file.path(tidy_root, run_label, paste0(run_label, "_gc_peak_heights.csv"))

file.exists(peaks_file)
ph <- readr::read_csv(peaks_file, show_col_types = FALSE)
glimpse(ph)
count(ph, chamber)
```

## QC - Diagnose and Flag Bad STD Injections

```{r qc-chunk}
stds_file <- file.path(tidy_root, run_label, paste0(run_label, "_gc_standards.csv"))

file.exists(stds_file)
stds <- readr::read_csv(stds_file, show_col_types = FALSE)

stds %>%
  pivot_longer(c(co2_height, ch4_height, n2o_height),
               names_to = "gas",
               values_to = "height") %>%
  ggplot(aes(row, height)) +
  geom_point() +
  facet_wrap(~ gas, scales = "free_y") 
```


## 2 - Convert peak heights to ppm

```{r run-step-2}
# Converts the one run we just produced
res_ppm <- heights_to_ppm(peaks_file)
```

```{r inspect-outputs}
std_file <- file.path(dirname(peaks_file), paste0(run_label, "_std_heights.csv"))
ppm_file <- file.path(dirname(peaks_file), paste0(run_label, "_ghg_ppm.csv"))

file.exists(std_file)
if (file.exists(std_file)) {
  std <- readr::read_csv(std_file, show_col_types = FALSE)
  print(std)
}

file.exists(ppm_file)
if (file.exists(ppm_file)) {
  ppm <- readr::read_csv(ppm_file, show_col_types = FALSE)
  glimpse(ppm)
  head(ppm, 10)
} else {
  message("No ghg_ppm.csv was written — likely because this run contains only standards (STD) and no non-STD samples.")
}
```




