---
title: "ppm-to-fluxes"
author: "Gavin McNicol"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-libraries}
library(tidyverse)
library(lubridate)
library(janitor)
library(stringi)
library(broom)
library(reshape2)
source("functions/convert_ppm_min_flux.R")
```

```{r}
files <- list.files("data/tidy/230720", pattern = "ppm")
ghg_ppm <- lapply(paste0("data/tidy/230720/", files), read_csv) %>% bind_rows()
ghg_ppm
```

```{r}
min_elapsed <- read_csv("data/tidy/230720/230720_min_elapsed.csv")
ghg_ppm <- ghg_ppm %>% 
  mutate(ecosystem = case_when(
    grepl("P", rep) ~ paste0(ecosystem, "P"),
    TRUE ~ ecosystem
  )) %>% 
  mutate(rep = str_remove(rep, "P")) %>% 
  separate(rep, c("pref", "suff"), "-") %>% 
  mutate(pref = as.numeric(pref),
         suff = as.numeric(suff)) %>% 
  left_join(min_elapsed, by = c("ecosystem", "pref", "suff", "timepoint")) %>% 
  mutate(date_collected = "230720",
         ecosystem = case_when(
           is.na(pref) ~ paste0(ecosystem, "P"),
           TRUE ~ ecosystem
         ))
ghg_ppm
```
### CO2 - Prairie

visualize linear co2 fits

```{r}
ghg_ppm %>% 
  filter(ecosystem == "Pr") %>% 
  ggplot(aes(x = min_elapsed, y = co2_ppm)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  facet_wrap(~ecosystem+pref+suff, 
             scales = "fixed") + 
  theme_classic() +
  theme(
  strip.background = element_blank(),
  strip.text.x = element_blank()
)
ggsave("data/diagnostics/230720_prairie_co2_lm.png")
```

### CO2 - North Woodland 

visualize linear co2 fits

```{r}
ghg_ppm %>% 
  filter(ecosystem == "NWd") %>% 
  
  ## Code to filter out bad time points
  mutate(co2_ppm = case_when(
    pref == 2 & suff == 3 & timepoint %in% c("t3", "t4") ~ NA,
    pref == 3 & suff %in% c(1,2) & timepoint %in% c("t3", "t4") ~ NA,
    TRUE ~ co2_ppm
  )) %>%
  
  ggplot(aes(x = min_elapsed, y = co2_ppm)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  facet_wrap(~ecosystem+pref+suff, 
             scales = "fixed") + 
  theme_classic() +
  theme(
  strip.background = element_blank(),
  strip.text.x = element_blank()
)
ggsave("data/diagnostics/230720_north_woodland_co2_lm.png")
```

### CO2 - South Woodland 

visualize linear co2 fits

```{r}
ghg_ppm %>% 
  filter(ecosystem == "SWd") %>% 
  
  ## Code to filter out bad time points\
  mutate(co2_ppm = case_when(
    pref == 3 & suff %in% c(3) & timepoint == "t4" ~ NA,
    TRUE ~ co2_ppm
  )) %>%
  
  ggplot(aes(x = min_elapsed, y = co2_ppm)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  facet_wrap(~ecosystem+pref+suff, 
             scales = "fixed") + 
  theme_classic() +
  theme(
  strip.background = element_blank(),
  strip.text.x = element_blank()
)
ggsave("data/diagnostics/230720_south_woodland_co2_lm.png")
```

### CO2 - Wetlands 

visualize linear co2 fits

```{r}
ghg_ppm %>% 
  filter(ecosystem == "Wt") %>% 
  
  # ## Code to filter out bad time points\
  # mutate(co2_ppm = case_when(
  #   pref == 2 & suff %in% c(4) & timepoint %in% c("t1", "t3") ~ NA,
  #   TRUE ~ co2_ppm
  # )) %>%
  
  ggplot(aes(x = min_elapsed, y = co2_ppm)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  facet_wrap(~ecosystem+pref+suff, 
             scales = "fixed") + 
  theme_classic() +
  theme(
  strip.background = element_blank(),
  strip.text.x = element_blank()
)
ggsave("data/diagnostics/230720_wetland_co2_lm.png")
```
fit linear co2 models

```{r}
ghg_ppm %>%
    filter(ecosystem %in% c("Pr", "NWd", "SWd", "Wt")) %>% 
    ## Code to filter out bad time points
  mutate(
    co2_ppm = case_when(
    ecosystem == "NWd" & pref == 3 & suff %in% c(1, 2) & timepoint == "t3" ~ NA,
    TRUE ~ co2_ppm),
  ) %>% 
  nest_by(ecosystem, pref, suff) %>% 
  mutate(lm_co2 = list(lm(co2_ppm ~ min_elapsed, data = data)),
         lm_ch4 = list(lm(ch4_ppm ~ min_elapsed, data = data)),
         lm_n2o = list(lm(n2o_ppm ~ min_elapsed, data = data))) %>%
  summarize(tidy(lm_co2)) %>% 
  filter(term == "min_elapsed") %>% 
  write_csv("data/final/230720_co2_fluxes.csv")
```


### CH4 - Prairie

visualize linear ch4 fits

```{r}
ghg_ppm %>% 
  filter(ecosystem == "Pr") %>% 
  ggplot(aes(x = min_elapsed, y = ch4_ppm)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  facet_wrap(~ecosystem+pref+suff, 
             scales = "fixed") + 
  theme_classic() +
  theme(
  strip.background = element_blank(),
  strip.text.x = element_blank()
)
ggsave("data/diagnostics/230720_prairie_ch4_lm.png")
```

### CH4 - North Woodland

visualize linear ch4 fits

```{r}
ghg_ppm %>% 
  filter(ecosystem == "NWd") %>% 
  
  # ## Code to filter out bad time points\
  # mutate(ch4_ppm = case_when(
  #   pref == 1 & suff %in% c(2, 3, 4) & timepoint == "t3" ~ NA,
  #   pref == 2 & suff %in% c(2, 3) & timepoint == "t3" ~ NA,
  #   pref == 3 & suff %in% c(1, 2) & timepoint == "t3" ~ NA,
  #   TRUE ~ ch4_ppm
  # )) %>% 
  
  ggplot(aes(x = min_elapsed, y = ch4_ppm)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  facet_wrap(~ecosystem+pref+suff, 
             scales = "fixed") + 
  theme_classic() +
  theme(
  strip.background = element_blank(),
  strip.text.x = element_blank()
)
ggsave("data/diagnostics/230720_north_woodland_ch4_lm.png")
```

### CH4 - South Woodland

visualize linear ch4 fits

```{r}
ghg_ppm %>% 
  filter(ecosystem == "SWd") %>% 
  
  # ## Code to filter out bad time points\
  # mutate(ch4_ppm = case_when(
  #   pref == 3 & suff %in% c(2) & timepoint == "t4" ~ NA,
  #   TRUE ~ ch4_ppm
  # )) %>%
  
  ggplot(aes(x = min_elapsed, y = ch4_ppm)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  facet_wrap(~ecosystem+pref+suff, 
             scales = "fixed") + 
  theme_classic() +
  theme(
  strip.background = element_blank(),
  strip.text.x = element_blank()
)
ggsave("data/diagnostics/230720_south_woodland_ch4_lm.png")
```

### CH4 - Wetlands 

visualize linear ch4 fits

```{r}
ghg_ppm %>% 
  filter(ecosystem == "Wt") %>% 
  
  # ## Code to filter out bad time points\
  # mutate(co2_ppm = case_when(
  #   pref == 2 & suff %in% c(4) & timepoint %in% c("t1", "t3") ~ NA,
  #   TRUE ~ co2_ppm
  # )) %>%
  
  ggplot(aes(x = min_elapsed, y = ch4_ppm)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  facet_wrap(~ecosystem+pref+suff, 
             scales = "fixed") + 
  theme_classic() +
  theme(
  strip.background = element_blank(),
  strip.text.x = element_blank()
)
ggsave("data/diagnostics/230720_wetland_ch4_lm.png")
```


fit linear ch4 models

```{r}
ghg_ppm %>%
    filter(ecosystem %in% c("Pr", "NWd", "SWd", "Wt")) %>% 
  ## Code to filter out bad time points
  mutate(ch4_ppm = case_when(
    ecosystem == "NWd" & pref == 1 & suff %in% c(2, 3, 4) & timepoint == "t3" ~ NA,
    ecosystem == "NWd" & pref == 2 & suff %in% c(2, 3) & timepoint == "t3" ~ NA,
    ecosystem == "NWd" & pref == 3 & suff %in% c(1, 2) & timepoint == "t3" ~ NA,
    ecosystem == "SWd" & pref == 3 & suff %in% c(2) & timepoint == "t4" ~ NA,
    TRUE ~ ch4_ppm
  )) %>% 
  nest_by(ecosystem, pref, suff) %>% 
  mutate(lm_co2 = list(lm(co2_ppm ~ min_elapsed, data = data)),
         lm_ch4 = list(lm(ch4_ppm ~ min_elapsed, data = data)),
         lm_n2o = list(lm(n2o_ppm ~ min_elapsed, data = data))) %>%
  summarize(tidy(lm_ch4)) %>% 
  filter(term == "min_elapsed") %>% 
  write_csv("data/final/230720_ch4_fluxes.csv")
```

### N2O - Prairie

visualize linear n2o fits

```{r}
ghg_ppm %>% 
  filter(ecosystem == "Pr") %>% 
  ggplot(aes(x = min_elapsed, y = n2o_ppm)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  facet_wrap(~ecosystem+pref+suff, 
             scales = "fixed") + 
  theme_classic() +
  theme(
  strip.background = element_blank(),
  strip.text.x = element_blank()
)
ggsave("data/diagnostics/230720_prairie_n2o_lm.png")
```

### N2O - North Woodland

visualize linear n2o fits

```{r}
ghg_ppm %>% 
  filter(ecosystem == "NWd") %>% 
  
  # ## Code to filter out bad time points\
  # mutate(n2o_ppm = case_when(
  #   pref == 1 & suff %in% c(2, 3, 4) & timepoint == "t3" ~ NA,
  #   pref == 2 & suff %in% c(2, 3) & timepoint == "t3" ~ NA,
  #   pref == 3 & suff %in% c(1, 2) & timepoint == "t3" ~ NA,
  #   TRUE ~ n2o_ppm
  # )) %>% 
  
  ggplot(aes(x = min_elapsed, y = n2o_ppm)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  facet_wrap(~ecosystem+pref+suff, 
             scales = "fixed") + 
  theme_classic() +
  theme(
  strip.background = element_blank(),
  strip.text.x = element_blank()
)
ggsave("data/diagnostics/230720_north_woodland_n2o_lm.png")
```

### N2O - South Woodland

visualize linear n2o fits

```{r}
ghg_ppm %>% 
  filter(ecosystem == "SWd") %>% 
  
  # ## Code to filter out bad time points\
  # mutate(n2o_ppm = case_when(
  #   pref == 3 & suff %in% c(2) & timepoint == "t4" ~ NA,
  #   TRUE ~ n2o_ppm
  # )) %>%
  
  ggplot(aes(x = min_elapsed, y = n2o_ppm)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  facet_wrap(~ecosystem+pref+suff, 
             scales = "fixed") + 
  theme_classic() +
  theme(
  strip.background = element_blank(),
  strip.text.x = element_blank()
)
ggsave("data/diagnostics/230720_south_woodland_n2o_lm.png")
```

### N2O - Wetlands 

visualize linear n2o fits

```{r}
ghg_ppm %>% 
  filter(ecosystem == "Wt") %>% 
  
  # ## Code to filter out bad time points\
  # mutate(co2_ppm = case_when(
  #   pref == 2 & suff %in% c(4) & timepoint %in% c("t1", "t3") ~ NA,
  #   TRUE ~ co2_ppm
  # )) %>%
  
  ggplot(aes(x = min_elapsed, y = n2o_ppm)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  facet_wrap(~ecosystem+pref+suff, 
             scales = "fixed") + 
  theme_classic() +
  theme(
  strip.background = element_blank(),
  strip.text.x = element_blank()
)
ggsave("data/diagnostics/230720_wetland_n2o_lm.png")
```


fit linear n2o models

```{r}
ghg_ppm %>%
    filter(ecosystem %in% c("Pr", "NWd", "SWd", "Wt")) %>% 
  ## Code to filter out bad time points\
  mutate(n2o_ppm = case_when(
    ecosystem == "NWd" & pref == 2 & suff == 3 & timepoint %in% c("t3", "t4") ~ NA,
    ecosystem == "NWd" & pref == 3 & suff %in% c(1,2) & timepoint %in% c("t3", "t4") ~ NA,
    ecosystem == "SWd" & pref == 3 & suff %in% c(3) & timepoint == "t4" ~ NA,
    TRUE ~ n2o_ppm
  )) %>% 
  nest_by(ecosystem, pref, suff) %>% 
  mutate(lm_co2 = list(lm(co2_ppm ~ min_elapsed, data = data)),
         lm_ch4 = list(lm(ch4_ppm ~ min_elapsed, data = data)),
         lm_n2o = list(lm(n2o_ppm ~ min_elapsed, data = data))) %>%
  summarize(tidy(lm_n2o)) %>% 
  filter(term == "min_elapsed") %>% 
  write_csv("data/final/230720_n2o_fluxes.csv")
```

## get fluxes to plot

```{r read}
co2 <- read_csv("data/final/230720_co2_fluxes.csv") %>% 
  mutate(gas = "co2")
ch4 <- read_csv("data/final/230720_ch4_fluxes.csv") %>% 
  mutate(gas = "ch4")
n2o <- read_csv("data/final/230720_n2o_fluxes.csv") %>% 
  mutate(gas = "n2o")
```

get chamber temps and vols

```{r chamber-temps}
chamber_temps <- read_csv("data/tidy/230720/230720_chamber_temps.csv")
chamber_vols <- read_csv("data/tidy/2023_chamber_vols.csv")
```


merge and convert units

```{r merge}
fluxes <- co2 %>% bind_rows(ch4) %>% bind_rows(n2o) %>% 
  select(ecosystem, pref, suff, gas, estimate, p.value) %>% 
  left_join(chamber_temps, by = c("ecosystem", "pref", "suff")) %>% 
  select(-temp_start, -temp_end) %>% 
  left_join(chamber_vols, by = c("ecosystem", "pref", "suff")) %>% 
  rowwise() %>% 
  mutate(flux = convert_ppm_min_flux(gas, estimate, total_vol_m3, average_temp_kelvin)) %>% 
  mutate(units = ifelse(gas == "co2", "umol_m2_s1", NA),
         units = ifelse(gas == "ch4", "nmol_m2_s1", units),
         units = ifelse(gas == "n2o", "nmol_m2_s1", units))
fluxes
```

save final flux data

```{r}
write_csv(fluxes, "data/final/230720_fluxes.csv")
```

### CO2 - Prairie Photosynthesis

visualize linear co2 fits DARK respiration

```{r}
ghg_ppm %>% 
  filter(ecosystem %in% c("PrP", "NWdP", "SWdP", "WtP") & suff == 2) %>% 
  # filter(ecosystem == "NWdP") %>% 
  # arrange(pref, suff)

   ## Code to filter out bad time points
  mutate(co2_ppm = case_when(
    timepoint == "t3" ~ NA,
    TRUE ~ co2_ppm
  )) %>%
  
  ggplot(aes(x = min_elapsed, y = co2_ppm)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  facet_wrap(~ecosystem+pref+suff, 
             scales = "free_y") + 
  theme_classic() +
  theme(
  strip.background = element_blank(),
  # strip.text.x = element_blank()
) + labs(title = "Dark Respiration")
ggsave("data/diagnostics/230720_prairie_co2p_lm_dark.png")
```

visualize linear co2 fits net CO2 flux

```{r}
ghg_ppm %>% 
  filter(ecosystem %in% c("PrP", "NWdP", "SWdP", "WtP") & suff == 1) %>% 
  # filter(ecosystem == "NWdP") %>% 
  # arrange(pref, suff)

   ## Code to filter out bad time points
  mutate(co2_ppm = case_when(
    timepoint == "t3" ~ NA,
    TRUE ~ co2_ppm
  )) %>%
  
  ggplot(aes(x = min_elapsed, y = co2_ppm)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  facet_wrap(~ecosystem+pref+suff, 
             scales = "free_y") + 
  theme_classic() +
  theme(
  strip.background = element_blank(),
  # strip.text.x = element_blank()
) + labs(title = "Dark Respiration")
ggsave("data/diagnostics/230720_prairie_co2p_lm_dark.png")
```

fit linear co2 models

```{r}
ghg_ppm %>%
    filter(ecosystem %in% c("PrP", "NWdP", "SWdP", "WtP")) %>% 
    ## Code to filter out bad time points
  mutate(
    co2_ppm = case_when(
    timepoint == "t3" ~ NA,
    TRUE ~ co2_ppm),
  ) %>% 
  nest_by(ecosystem, pref, suff) %>% 
  mutate(lm_co2 = list(lm(co2_ppm ~ min_elapsed, data = data))) %>%
  summarize(tidy(lm_co2)) %>% 
  filter(term == "min_elapsed") %>% 
  write_csv("data/final/230720_co2p_fluxes.csv")
```

get chamber temps and vols

```{r chamber-temps}
chamber_temps <- read_csv("data/tidy/230720/230720_chamber_temps.csv")
chamber_vols <- read_csv("data/tidy/2023_chamber_vols.csv")
```

## get fluxes to plot

```{r read}
co2 <- read_csv("data/final/230720_co2p_fluxes.csv") %>% 
  mutate(gas = "co2")
```

merge and convert units

```{r merge}
fluxes <- co2 %>% 
  select(ecosystem, pref, suff, gas, estimate, p.value) %>% 
  left_join(chamber_temps, by = c("ecosystem", "pref", "suff")) %>% 
  select(-temp_start, -temp_end, -date) %>% 
  left_join(chamber_vols, by = c("ecosystem", "pref", "suff")) %>% 
  rowwise() %>% 
  mutate(flux = convert_ppm_min_flux(gas, estimate, total_vol_m3, average_temp_kelvin)) %>% 
  mutate(units = ifelse(gas == "co2", "umol_m2_s1", NA))
fluxes
```

save final flux data

```{r}
write_csv(fluxes, "data/final/230720_photosynthesis.csv")
```