---
title: "Gas Fluxes: GC Peak Height Extraction"
author: "Gavin"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Purpose: Process GC data to get soil chamber GHG fluxes from JWP

Data:
1. Find data
  - .txt files within the directory


```{r}
library(tidyverse)
library(lubridate)
library(janitor)
library(stringi)
library(broom)
library(reshape2)
```

## TIMEPOINTS

```{r get-sampling-dates}
# dates
```

list the 123 .txt files to read

## t0

```{r list-files}
files <- list.files("data/gc-raw/20230720/t0", pattern = ".txt")
length(files)
```

get the sample log

**To get the sample log I need to export a subset of the full .xlsx sample log data as a .csv file**

```{r read-sample-log}
sample_log <- read_csv("data/gc-raw/20230720/t0/sample_log.csv", na = c("N/A", "<NA>")) %>% 
  janitor::clean_names() %>% 
  mutate(date_collected = mdy(date_collected),
         gc_run_date = mdy(gc_run_date)) %>% 
  mutate(row = as.numeric(substr(filename, 13, 15)),
         timepoint = "t0")

glimpse(sample_log)
```

read in gc files

```{r read-gc-files, output = F, echo = F}
data <- lapply(paste0("data/gc-raw/20230720/t0/", files), read_fwf)
data[[39]]
```

get CO2 peak heights

```{r co2-heights}
co2_height_1 <- c()
co2_height_2 <- c()
co2_height_3 <- c()
for(i in 1:length(files)){
  co2_height_1 <- as.numeric(append(co2_height_1, strsplit(data[[i]][84,]$X1, ",")[[1]][6]))
  co2_height_2 <- as.numeric(append(co2_height_2, strsplit(data[[i]][84,]$X1, ",")[[1]][6]))
  co2_height_3 <- as.numeric(append(co2_height_3, strsplit(data[[i]][85,]$X1, ",")[[1]][6])) 
}

co2_height <- bind_cols(co2_height_1, co2_height_2, co2_height_3) %>% 
  rename("col1" = `...1`,
         "col2" = `...2`,
         "col3" = `...3`) %>% 
  mutate(col1 = ifelse(col1 < 100000, NA, col1),
         col2 = ifelse(col2 < 950000 | col2 > 105000, NA, col2),
         col3 = ifelse(col3 < 100000, NA, col3),
         co2_height = coalesce(col1, col2, col3)) %>% 
  pull(co2_height)
co2_height
```


```{r ch4-heights}
ch4_height_1 <- c()
ch4_height_2 <- c()
ch4_height_3 <- c()
for(i in 1:length(files)){
  ch4_height_1 <- as.numeric(append(ch4_height_1, strsplit(data[[i]][83,]$X1, ",")[[1]][6]))
  ch4_height_2 <- as.numeric(append(ch4_height_2, strsplit(data[[i]][83,]$X1, ",")[[1]][6]))
  ch4_height_3 <- as.numeric(append(ch4_height_3, strsplit(data[[i]][84,]$X1, ",")[[1]][6]))
}

ch4_height <- bind_cols(ch4_height_1, ch4_height_2, ch4_height_3) %>% 
  rename("col1" = `...1`,
         "col2" = `...2`,
         "col3" = `...3`) %>% 
  mutate(col1 = ifelse(col1 < 1000 | col1 > 3000, NA, col1),
         col2 = ifelse(col2 < 9500 | col2 > 11500, NA, col2),
         col3 = ifelse(col3 < 1000 | col3 > 3000, NA, col3)) %>% 
  mutate(ch4_height = coalesce(col1, col2, col3)) %>% 
  pull(ch4_height)
ch4_height
```

```{r n2o-heigts}
n2o_height_2 <- c()
n2o_height_3 <- c()
n2o_height_4 <- c()
n2o_height_5 <- c()
for(i in 1:length(files)){
    n2o_height_2 <- as.numeric(append(n2o_height_2, strsplit(data[[i]][93,]$X1, ",")[[1]][6]))
    n2o_height_3 <- as.numeric(append(n2o_height_3, strsplit(data[[i]][94,]$X1, ",")[[1]][6]))
    n2o_height_4 <- as.numeric(append(n2o_height_4, strsplit(data[[i]][95,]$X1, ",")[[1]][6]))
    n2o_height_5 <- as.numeric(append(n2o_height_5, strsplit(data[[i]][96,]$X1, ",")[[1]][6]))

}

n2o_height <- bind_cols(n2o_height_2, n2o_height_3, n2o_height_4, n2o_height_5) %>% 
  rename("col1" = `...1`,
         "col2" = `...2`,
         "col3" = `...3`,
         "col4" = `...4`) %>% 
  mutate(col1 = ifelse(col1 < 2000, NA, col1),
         col2 = ifelse(col2 < 2000, NA, col2),
         col3 = ifelse(col3 < 2000, NA, col3),
         col4 = ifelse(col4 < 2000, NA, col4)) %>% 
  mutate(n2o_height = coalesce(col1, col2, col3, col4)) %>% 
  pull(n2o_height)
n2o_height
```

bind peak heights

```{r}
peak_heights <- bind_cols(1:length(co2_height), co2_height, ch4_height, n2o_height) %>% 
  rename("row" = `...1`, "co2_height" = `...2`, "ch4_height" = `...3`, "n2o_height" = `...4`)
```

bind sample log to peak heights
```{r}
gc_data_tidy <- left_join(peak_heights, sample_log, by = "row") %>% 
  mutate(ecosystem = ifelse(is.na(ecosystem), "STD", ecosystem)) %>% 
  select(row, date_collected, timepoint, gc_run_date, filename, ecosystem, rep = chamber_number, 2, 3, 4, 5)
gc_data_tidy
```

```{r}
write_csv(gc_data_tidy, "data/tidy/230720/230720_t0_gc_peak_heights.csv")
```

## Re-read in tidy data

```{r}
gc_data_tidy <- read_csv("data/tidy/230720/230720_t0_gc_peak_heights.csv")
```


get std stats across all runs

```{r}
std_heights <- gc_data_tidy %>% 
  filter(ecosystem == "STD" & !is.na(gc_run_date) & n2o_height > 10000 & n2o_height < 15000,
         row != 1) %>% 
  group_by(gc_run_date) %>% 
  summarize(mean_co2_height = mean(co2_height, na.rm = T),
            mean_ch4_height = mean(ch4_height, na.rm = T),
            mean_n2o_height = mean(n2o_height, na.rm = T),
            
            co2_precision = sd(co2_height, na.rm = T)/mean_co2_height*100,
            ch4_precision = sd(ch4_height, na.rm = T)/mean_ch4_height*100,
            n2o_precision = sd(n2o_height, na.rm = T)/mean_n2o_height*100) 
std_heights
```

combine std heights with unknowns

```{r}
unknown_std_heights <- gc_data_tidy %>% 
  filter(ecosystem != "STD") %>% 
  left_join(std_heights, by = "gc_run_date")
```

calculate ghg in ppm

```{r}
ghg_ppm <- unknown_std_heights %>% 
  mutate(co2_ppm = co2_height * (998/mean_co2_height),
         ch4_ppm = ch4_height * (10.2/mean_ch4_height),
         n2o_ppm = n2o_height * (1/mean_n2o_height)) 
ghg_ppm
```

write data

```{r}
write_csv(ghg_ppm, "data/tidy/230720/230720_t0_ghg_ppm.csv")
```


## t1

```{r list-files}
files <- list.files("data/gc-raw/20230720/t1", pattern = ".txt")
length(files)
```

get the sample log

**To get the sample log I need to export a subset of the full .xlsx sample log data as a .csv file**

```{r read-sample-log}
sample_log <- read_csv("data/gc-raw/20230720/t1/sample_log.csv", na = c("N/A", "<NA>")) %>% 
  janitor::clean_names() %>% 
  mutate(date_collected = mdy(date_collected),
         gc_run_date = mdy(gc_run_date)) %>% 
  mutate(row = as.numeric(substr(filename, 13, 15)),
         timepoint = "t1",
         row = 1:n())

glimpse(sample_log)
```

read in gc files

```{r read-gc-files, output = F, echo = F}
data <- lapply(paste0("data/gc-raw/20230720/t1/", files), read_fwf)
data[[98]]
```

get CO2 peak heights

```{r co2-heights}
co2_height_1 <- c()
co2_height_2 <- c()
co2_height_3 <- c()
for(i in 1:length(files)){
  co2_height_1 <- as.numeric(append(co2_height_1, strsplit(data[[i]][84,]$X1, ",")[[1]][6]))
  co2_height_2 <- as.numeric(append(co2_height_2, strsplit(data[[i]][84,]$X1, ",")[[1]][6]))
  co2_height_3 <- as.numeric(append(co2_height_3, strsplit(data[[i]][85,]$X1, ",")[[1]][6])) 
}

co2_height <- bind_cols(co2_height_1, co2_height_2, co2_height_3) %>% 
  rename("col1" = `...1`,
         "col2" = `...2`,
         "col3" = `...3`) %>% 
  mutate(col1 = ifelse(col1 < 100000, NA, col1),
         col2 = ifelse(col2 < 950000 | col2 > 105000, NA, col2),
         col3 = ifelse(col3 < 100000, NA, col3),
         co2_height = coalesce(col1, col2, col3)) %>% 
  pull(co2_height)
co2_height
```


```{r ch4-heights}
ch4_height_1 <- c()
ch4_height_2 <- c()
ch4_height_3 <- c()
for(i in 1:length(files)){
  ch4_height_1 <- as.numeric(append(ch4_height_1, strsplit(data[[i]][83,]$X1, ",")[[1]][6]))
  ch4_height_2 <- as.numeric(append(ch4_height_2, strsplit(data[[i]][83,]$X1, ",")[[1]][6]))
  ch4_height_3 <- as.numeric(append(ch4_height_3, strsplit(data[[i]][84,]$X1, ",")[[1]][6]))
}

ch4_height <- bind_cols(ch4_height_1, ch4_height_2, ch4_height_3) %>% 
  rename("col1" = `...1`,
         "col2" = `...2`,
         "col3" = `...3`) %>% 
  mutate(col1 = ifelse(col1 < 1000 | col1 > 3000, NA, col1),
         col2 = ifelse(col2 < 9500 | col2 > 11500, NA, col2),
         col3 = ifelse(col3 < 1000 | col3 > 3000, NA, col3)) %>% 
  mutate(ch4_height = coalesce(col1, col2, col3)) %>% 
  pull(ch4_height)
ch4_height
```

```{r n2o-heigts}
n2o_height_2 <- c()
n2o_height_3 <- c()
n2o_height_4 <- c()
n2o_height_5 <- c()
for(i in 1:length(files)){
    n2o_height_2 <- as.numeric(append(n2o_height_2, strsplit(data[[i]][93,]$X1, ",")[[1]][6]))
    n2o_height_3 <- as.numeric(append(n2o_height_3, strsplit(data[[i]][94,]$X1, ",")[[1]][6]))
    n2o_height_4 <- as.numeric(append(n2o_height_4, strsplit(data[[i]][95,]$X1, ",")[[1]][6]))
    n2o_height_5 <- as.numeric(append(n2o_height_5, strsplit(data[[i]][96,]$X1, ",")[[1]][6]))

}

n2o_height <- bind_cols(n2o_height_2, n2o_height_3, n2o_height_4, n2o_height_5) %>% 
  rename("col1" = `...1`,
         "col2" = `...2`,
         "col3" = `...3`,
         "col4" = `...4`) %>% 
  mutate(col1 = ifelse(col1 < 2000, NA, col1),
         col2 = ifelse(col2 < 2000, NA, col2),
         col3 = ifelse(col3 < 2000, NA, col3),
         col4 = ifelse(col4 < 2000, NA, col4)) %>% 
  mutate(n2o_height = coalesce(col1, col2, col3, col4)) %>% 
  pull(n2o_height)
n2o_height
```

bind peak heights

```{r}
peak_heights <- bind_cols(1:length(co2_height), co2_height, ch4_height, n2o_height) %>% 
  rename("row" = `...1`, "co2_height" = `...2`, "ch4_height" = `...3`, "n2o_height" = `...4`)
```

bind sample log to peak heights
```{r}
gc_data_tidy <- left_join(peak_heights, sample_log, by = "row") %>% 
  mutate(ecosystem = ifelse(is.na(ecosystem), "STD", ecosystem)) %>% 
  select(row, date_collected, timepoint, gc_run_date, filename, ecosystem, rep = chamber_number, 2, 3, 4, 5)
gc_data_tidy
```

```{r}
write_csv(gc_data_tidy, "data/tidy/230720/230720_t1_gc_peak_heights.csv")
```

## Re-read in tidy data

```{r}
gc_data_tidy <- read_csv("data/tidy/230720/230720_t1_gc_peak_heights.csv")
```


get std stats across all runs

```{r}
std_heights <- gc_data_tidy %>% 
  filter(ecosystem == "STD" & !is.na(gc_run_date) & n2o_height > 10000,
         row != 1) %>% 
  group_by(gc_run_date) %>% 
  summarize(mean_co2_height = mean(co2_height, na.rm = T),
            mean_ch4_height = mean(ch4_height, na.rm = T),
            mean_n2o_height = mean(n2o_height, na.rm = T),
            
            co2_precision = sd(co2_height, na.rm = T)/mean_co2_height*100,
            ch4_precision = sd(ch4_height, na.rm = T)/mean_ch4_height*100,
            n2o_precision = sd(n2o_height, na.rm = T)/mean_n2o_height*100) 
std_heights
```

combine std heights with unknowns

```{r}
unknown_std_heights <- gc_data_tidy %>% 
  filter(ecosystem != "STD") %>% 
  left_join(std_heights, by = "gc_run_date")
```

calculate ghg in ppm

```{r}
ghg_ppm <- unknown_std_heights %>% 
  mutate(co2_ppm = co2_height * (998/mean_co2_height),
         ch4_ppm = ch4_height * (10.2/mean_ch4_height),
         n2o_ppm = n2o_height * (1/mean_n2o_height)) 
ghg_ppm
```

write data

```{r}
write_csv(ghg_ppm, "data/tidy/230720/230720_t1_ghg_ppm.csv")
```

## t2

```{r list-files}
files <- list.files("data/gc-raw/20230720/t2", pattern = ".txt")
length(files)
```

get the sample log

**To get the sample log I need to export a subset of the full .xlsx sample log data as a .csv file**

```{r read-sample-log}
sample_log <- read_csv("data/gc-raw/20230720/t2/sample_log.csv", na = c("N/A", "<NA>")) %>% 
  janitor::clean_names() %>% 
  mutate(date_collected = mdy(date_collected),
         gc_run_date = mdy(gc_run_date)) %>% 
  mutate(row = as.numeric(substr(filename, 13, 15)),
         timepoint = "t2",
         row = 1:n())

glimpse(sample_log)
```

read in gc files

```{r read-gc-files, output = F, echo = F}
data <- lapply(paste0("data/gc-raw/20230720/t2/", files), read_fwf)
data[[80]]
```

get CO2 peak heights

```{r co2-heights}
co2_height_1 <- c()
co2_height_2 <- c()
co2_height_3 <- c()
for(i in 1:length(files)){
  co2_height_1 <- as.numeric(append(co2_height_1, strsplit(data[[i]][84,]$X1, "\t")[[1]][6]))
  co2_height_2 <- as.numeric(append(co2_height_2, strsplit(data[[i]][84,]$X1, "\t")[[1]][6]))
  co2_height_3 <- as.numeric(append(co2_height_3, strsplit(data[[i]][85,]$X1, "\t")[[1]][6])) 
}

co2_height <- bind_cols(co2_height_1, co2_height_2, co2_height_3) %>% 
  rename("col1" = `...1`,
         "col2" = `...2`,
         "col3" = `...3`) %>% 
  mutate(col1 = ifelse(col1 < 100000, NA, col1),
         col2 = ifelse(col2 < 950000 | col2 > 105000, NA, col2),
         col3 = ifelse(col3 < 100000, NA, col3),
         co2_height = coalesce(col1, col2, col3)) %>% 
  pull(co2_height)
co2_height
```


```{r ch4-heights}
ch4_height_1 <- c()
ch4_height_2 <- c()
ch4_height_3 <- c()
for(i in 1:length(files)){
  ch4_height_1 <- as.numeric(append(ch4_height_1, strsplit(data[[i]][83,]$X1, "\t")[[1]][6]))
  ch4_height_2 <- as.numeric(append(ch4_height_2, strsplit(data[[i]][83,]$X1, "\t")[[1]][6]))
  ch4_height_3 <- as.numeric(append(ch4_height_3, strsplit(data[[i]][84,]$X1, "\t")[[1]][6]))
}

ch4_height <- bind_cols(ch4_height_1, ch4_height_2, ch4_height_3) %>% 
  rename("col1" = `...1`,
         "col2" = `...2`,
         "col3" = `...3`) %>% 
  mutate(col1 = ifelse(col1 < 1000 | col1 > 3000, NA, col1),
         col2 = ifelse(col2 < 9500 | col2 > 11500, NA, col2),
         col3 = ifelse(col3 < 1000 | col3 > 3000, NA, col3)) %>% 
  mutate(ch4_height = coalesce(col1, col2, col3)) %>% 
  pull(ch4_height)
ch4_height
```

```{r n2o-heigts}
n2o_height_2 <- c()
n2o_height_3 <- c()
n2o_height_4 <- c()
n2o_height_5 <- c()
for(i in 1:length(files)){
    n2o_height_2 <- as.numeric(append(n2o_height_2, strsplit(data[[i]][93,]$X1, "\t")[[1]][6]))
    n2o_height_3 <- as.numeric(append(n2o_height_3, strsplit(data[[i]][94,]$X1, "\t")[[1]][6]))
    n2o_height_4 <- as.numeric(append(n2o_height_4, strsplit(data[[i]][95,]$X1, "\t")[[1]][6]))
    n2o_height_5 <- as.numeric(append(n2o_height_5, strsplit(data[[i]][96,]$X1, "\t")[[1]][6]))

}

n2o_height <- bind_cols(n2o_height_2, n2o_height_3, n2o_height_4, n2o_height_5) %>% 
  rename("col1" = `...1`,
         "col2" = `...2`,
         "col3" = `...3`,
         "col4" = `...4`) %>% 
  mutate(col1 = ifelse(col1 < 2000, NA, col1),
         col2 = ifelse(col2 < 2000, NA, col2),
         col3 = ifelse(col3 < 2000, NA, col3),
         col4 = ifelse(col4 < 2000, NA, col4)) %>% 
  mutate(n2o_height = coalesce(col1, col2, col3, col4)) %>% 
  pull(n2o_height)
n2o_height
```

bind peak heights

```{r}
peak_heights <- bind_cols(1:length(co2_height), co2_height, ch4_height, n2o_height) %>% 
  rename("row" = `...1`, "co2_height" = `...2`, "ch4_height" = `...3`, "n2o_height" = `...4`)
```

bind sample log to peak heights
```{r}
gc_data_tidy <- left_join(peak_heights, sample_log, by = "row") %>% 
  mutate(ecosystem = ifelse(is.na(ecosystem), "STD", ecosystem)) %>% 
  select(row, date_collected, timepoint, gc_run_date, filename, ecosystem, rep = chamber_number, 2, 3, 4, 5)
gc_data_tidy
```

```{r}
write_csv(gc_data_tidy, "data/tidy/230720/230720_t2_gc_peak_heights.csv")
```

## Re-read in tidy data

```{r}
gc_data_tidy <- read_csv("data/tidy/230720/230720_t2_gc_peak_heights.csv")
```


get std stats across all runs

```{r}
std_heights <- gc_data_tidy %>% 
  filter(ecosystem == "STD" & !is.na(gc_run_date) & n2o_height > 10000,
         row != 1) %>% 
  group_by(gc_run_date) %>% 
  summarize(mean_co2_height = mean(co2_height, na.rm = T),
            mean_ch4_height = mean(ch4_height, na.rm = T),
            mean_n2o_height = mean(n2o_height, na.rm = T),
            
            co2_precision = sd(co2_height, na.rm = T)/mean_co2_height*100,
            ch4_precision = sd(ch4_height, na.rm = T)/mean_ch4_height*100,
            n2o_precision = sd(n2o_height, na.rm = T)/mean_n2o_height*100) 
std_heights
```

combine std heights with unknowns

```{r}
unknown_std_heights <- gc_data_tidy %>% 
  filter(ecosystem != "STD") %>% 
  left_join(std_heights, by = "gc_run_date")
```

calculate ghg in ppm

```{r}
ghg_ppm <- unknown_std_heights %>% 
  mutate(co2_ppm = co2_height * (998/mean_co2_height),
         ch4_ppm = ch4_height * (10.2/mean_ch4_height),
         n2o_ppm = n2o_height * (1/mean_n2o_height)) 
ghg_ppm
```

write data

```{r}
write_csv(ghg_ppm, "data/tidy/230720/230720_t2_ghg_ppm.csv")
```


## t3

```{r list-files}
files <- list.files("data/gc-raw/20230720/t3", pattern = ".txt")
length(files)
```

get the sample log

**To get the sample log I need to export a subset of the full .xlsx sample log data as a .csv file**

```{r read-sample-log}
sample_log <- read_csv("data/gc-raw/20230720/t3/sample_log.csv", na = c("N/A", "<NA>")) %>% 
  janitor::clean_names() %>% 
  mutate(date_collected = mdy(date_collected),
         gc_run_date = mdy(gc_run_date)) %>% 
  mutate(row = as.numeric(substr(filename, 13, 15)),
         timepoint = "t3")

glimpse(sample_log)
```

read in gc files

```{r read-gc-files, output = F, echo = F}
data <- lapply(paste0("data/gc-raw/20230720/t3/", files), read_fwf)
data[[30]]
```

get CO2 peak heights

```{r co2-heights}
co2_height_1 <- c()
co2_height_2 <- c()
co2_height_3 <- c()
for(i in 1:length(files)){
  co2_height_1 <- as.numeric(append(co2_height_1, strsplit(data[[i]][84,]$X1, "\t")[[1]][6]))
  co2_height_2 <- as.numeric(append(co2_height_2, strsplit(data[[i]][84,]$X1, "\t")[[1]][6]))
  co2_height_3 <- as.numeric(append(co2_height_3, strsplit(data[[i]][85,]$X1, "\t")[[1]][6])) 
}

co2_height <- bind_cols(co2_height_1, co2_height_2, co2_height_3) %>% 
  rename("col1" = `...1`,
         "col2" = `...2`,
         "col3" = `...3`) %>% 
  mutate(col1 = ifelse(col1 < 100000, NA, col1),
         col2 = ifelse(col2 < 950000 | col2 > 105000, NA, col2),
         col3 = ifelse(col3 < 100000, NA, col3),
         co2_height = coalesce(col1, col2, col3)) %>% 
  pull(co2_height)
co2_height
```


```{r ch4-heights}
ch4_height_1 <- c()
ch4_height_2 <- c()
ch4_height_3 <- c()
for(i in 1:length(files)){
  ch4_height_1 <- as.numeric(append(ch4_height_1, strsplit(data[[i]][83,]$X1, "\t")[[1]][6]))
  ch4_height_2 <- as.numeric(append(ch4_height_2, strsplit(data[[i]][83,]$X1, "\t")[[1]][6]))
  ch4_height_3 <- as.numeric(append(ch4_height_3, strsplit(data[[i]][84,]$X1, "\t")[[1]][6]))
}

ch4_height <- bind_cols(ch4_height_1, ch4_height_2, ch4_height_3) %>% 
  rename("col1" = `...1`,
         "col2" = `...2`,
         "col3" = `...3`) %>% 
  mutate(col1 = ifelse(col1 < 1000 | col1 > 3000, NA, col1),
         col2 = ifelse(col2 < 9500 | col2 > 11500, NA, col2),
         col3 = ifelse(col3 < 1000 | col3 > 3000, NA, col3)) %>% 
  mutate(ch4_height = coalesce(col1, col2, col3)) %>% 
  pull(ch4_height)
ch4_height
```

```{r n2o-heigts}
n2o_height_2 <- c()
n2o_height_3 <- c()
n2o_height_4 <- c()
n2o_height_5 <- c()
for(i in 1:length(files)){
    n2o_height_2 <- as.numeric(append(n2o_height_2, strsplit(data[[i]][93,]$X1, "\t")[[1]][6]))
    n2o_height_3 <- as.numeric(append(n2o_height_3, strsplit(data[[i]][94,]$X1, "\t")[[1]][6]))
    n2o_height_4 <- as.numeric(append(n2o_height_4, strsplit(data[[i]][95,]$X1, "\t")[[1]][6]))
    n2o_height_5 <- as.numeric(append(n2o_height_5, strsplit(data[[i]][96,]$X1, "\t")[[1]][6]))

}

n2o_height <- bind_cols(n2o_height_2, n2o_height_3, n2o_height_4, n2o_height_5) %>% 
  rename("col1" = `...1`,
         "col2" = `...2`,
         "col3" = `...3`,
         "col4" = `...4`) %>% 
  mutate(col1 = ifelse(col1 < 4000, NA, col1),
         col2 = ifelse(col2 < 4000, NA, col2),
         col3 = ifelse(col3 < 2000, NA, col3),
         col4 = ifelse(col4 < 2000, NA, col4)) %>% 
  mutate(n2o_height = coalesce(col1, col2, col3, col4)) %>% 
  pull(n2o_height)
n2o_height
```

bind peak heights

```{r}
peak_heights <- bind_cols(1:length(co2_height), co2_height, ch4_height, n2o_height) %>% 
  rename("row" = `...1`, "co2_height" = `...2`, "ch4_height" = `...3`, "n2o_height" = `...4`)
```

bind sample log to peak heights
```{r}
gc_data_tidy <- left_join(peak_heights, sample_log, by = "row") %>% 
  mutate(ecosystem = ifelse(is.na(ecosystem), "STD", ecosystem)) %>% 
  select(row, date_collected, timepoint, gc_run_date, filename, ecosystem, rep = chamber_number, 2, 3, 4, 5)
gc_data_tidy
```

```{r}
write_csv(gc_data_tidy, "data/tidy/230720/230720_t3_gc_peak_heights.csv")
```

## Re-read in tidy data

```{r}
gc_data_tidy <- read_csv("data/tidy/230720/230720_t3_gc_peak_heights.csv")
```


get std stats across all runs

```{r}
std_heights <- gc_data_tidy %>% 
  filter(ecosystem == "STD" & !is.na(gc_run_date) & n2o_height > 10000,
         row != 1) %>% 
  group_by(gc_run_date) %>% 
  summarize(mean_co2_height = mean(co2_height, na.rm = T),
            mean_ch4_height = mean(ch4_height, na.rm = T),
            mean_n2o_height = mean(n2o_height, na.rm = T),
            
            co2_precision = sd(co2_height, na.rm = T)/mean_co2_height*100,
            ch4_precision = sd(ch4_height, na.rm = T)/mean_ch4_height*100,
            n2o_precision = sd(n2o_height, na.rm = T)/mean_n2o_height*100) 
std_heights
```

combine std heights with unknowns

```{r}
unknown_std_heights <- gc_data_tidy %>% 
  filter(ecosystem != "STD") %>% 
  left_join(std_heights, by = "gc_run_date")
```

calculate ghg in ppm

```{r}
ghg_ppm <- unknown_std_heights %>% 
  mutate(co2_ppm = co2_height * (998/mean_co2_height),
         ch4_ppm = ch4_height * (10.2/mean_ch4_height),
         n2o_ppm = n2o_height * (1/mean_n2o_height)) 
ghg_ppm
```

write data

```{r}
write_csv(ghg_ppm, "data/tidy/230720/230720_t3_ghg_ppm.csv")
```


## t4

```{r list-files}
files <- list.files("data/gc-raw/20230720/t4", pattern = ".txt")
length(files)
```

get the sample log

**To get the sample log I need to export a subset of the full .xlsx sample log data as a .csv file**

```{r read-sample-log}
sample_log <- read_csv("data/gc-raw/20230720/t4/sample_log.csv", na = c("N/A", "<NA>")) %>% 
  janitor::clean_names() %>% 
  mutate(date_collected = mdy(date_collected),
         gc_run_date = mdy(gc_run_date)) %>% 
  mutate(row = as.numeric(substr(filename, 13, 15)),
         timepoint = "t4")

glimpse(sample_log)
```

read in gc files

```{r read-gc-files, output = F, echo = F}
data <- lapply(paste0("data/gc-raw/20230720/t4/", files), read_fwf)
data[[39]]
```

get CO2 peak heights

```{r co2-heights}
co2_height_1 <- c()
co2_height_2 <- c()
co2_height_3 <- c()
for(i in 1:length(files)){
  co2_height_1 <- as.numeric(append(co2_height_1, strsplit(data[[i]][84,]$X1, "\t")[[1]][6]))
  co2_height_2 <- as.numeric(append(co2_height_2, strsplit(data[[i]][84,]$X1, "\t")[[1]][6]))
  co2_height_3 <- as.numeric(append(co2_height_3, strsplit(data[[i]][85,]$X1, "\t")[[1]][6])) 
}

co2_height <- bind_cols(co2_height_1, co2_height_2, co2_height_3) %>% 
  rename("col1" = `...1`,
         "col2" = `...2`,
         "col3" = `...3`) %>% 
  mutate(col1 = ifelse(col1 < 400000, NA, col1),
         col2 = ifelse(col2 < 950000 | col2 > 105000, NA, col2),
         col3 = ifelse(col3 < 400000, NA, col3),
         co2_height = coalesce(col1, col2, col3)) %>% 
  pull(co2_height)
co2_height
```


```{r ch4-heights}
ch4_height_1 <- c()
ch4_height_2 <- c()
ch4_height_3 <- c()
ch4_height_4 <- c()
for(i in 1:length(files)){
  ch4_height_1 <- as.numeric(append(ch4_height_1, strsplit(data[[i]][83,]$X1, "\t")[[1]][6]))
  ch4_height_2 <- as.numeric(append(ch4_height_2, strsplit(data[[i]][83,]$X1, "\t")[[1]][6]))
  ch4_height_3 <- as.numeric(append(ch4_height_3, strsplit(data[[i]][84,]$X1, "\t")[[1]][6]))
    ch4_height_4 <- as.numeric(append(ch4_height_4, strsplit(data[[i]][84,]$X1, "\t")[[1]][6]))
}

ch4_height <- bind_cols(ch4_height_1, ch4_height_2, ch4_height_3, ch4_height_4) %>% 
  rename("col1" = `...1`,
         "col2" = `...2`,
         "col3" = `...3`,
         "col4" = `...4`) %>% 
  mutate(col1 = ifelse(col1 < 1000 | col1 > 3000, NA, col1),
         col2 = ifelse(col2 < 9000 | col2 > 12000, NA, col2),
         col3 = ifelse(col3 < 1000 | col3 > 3000, NA, col3),
         col4 = ifelse(col4 < 9000 | col4 > 12000, NA, col4)) %>% 
  mutate(ch4_height = coalesce(col1, col2, col3, col4)) %>% 
  pull(ch4_height)
ch4_height
```

```{r n2o-heigts}
n2o_height_2 <- c()
n2o_height_3 <- c()
n2o_height_4 <- c()
n2o_height_5 <- c()
for(i in 1:length(files)){
    n2o_height_2 <- as.numeric(append(n2o_height_2, strsplit(data[[i]][93,]$X1, "\t")[[1]][6]))
    n2o_height_3 <- as.numeric(append(n2o_height_3, strsplit(data[[i]][94,]$X1, "\t")[[1]][6]))
    n2o_height_4 <- as.numeric(append(n2o_height_4, strsplit(data[[i]][95,]$X1, "\t")[[1]][6]))
    n2o_height_5 <- as.numeric(append(n2o_height_5, strsplit(data[[i]][96,]$X1, "\t")[[1]][6]))

}

n2o_height <- bind_cols(n2o_height_2, n2o_height_3, n2o_height_4, n2o_height_5) %>% 
  rename("col1" = `...1`,
         "col2" = `...2`,
         "col3" = `...3`,
         "col4" = `...4`) %>% 
  mutate(col1 = ifelse(col1 < 2000, NA, col1),
         col2 = ifelse(col2 < 2000, NA, col2),
         col3 = ifelse(col3 < 2000, NA, col3),
         col4 = ifelse(col4 < 2000, NA, col4)) %>% 
  mutate(n2o_height = coalesce(col1, col2, col3, col4)) %>% 
  pull(n2o_height)
n2o_height
```

bind peak heights

```{r}
peak_heights <- bind_cols(1:length(co2_height), co2_height, ch4_height, n2o_height) %>% 
  rename("row" = `...1`, "co2_height" = `...2`, "ch4_height" = `...3`, "n2o_height" = `...4`)
```

bind sample log to peak heights
```{r}
gc_data_tidy <- left_join(peak_heights, sample_log, by = "row") %>% 
  mutate(ecosystem = ifelse(is.na(ecosystem), "STD", ecosystem)) %>% 
  select(row, date_collected, timepoint, gc_run_date, filename, ecosystem, rep = chamber_number, 2, 3, 4, 5)
gc_data_tidy
```

```{r}
write_csv(gc_data_tidy, "data/tidy/230720/230720_t4_gc_peak_heights.csv")
```

## Re-read in tidy data

```{r}
gc_data_tidy <- read_csv("data/tidy/230720/230720_t4_gc_peak_heights.csv")
```


get std stats across all runs

```{r}
std_heights <- gc_data_tidy %>% 
  filter(ecosystem == "STD" & !is.na(gc_run_date) & n2o_height > 10000,
         row != 1) %>% 
  group_by(gc_run_date) %>% 
  summarize(mean_co2_height = mean(co2_height, na.rm = T),
            mean_ch4_height = mean(ch4_height, na.rm = T),
            mean_n2o_height = mean(n2o_height, na.rm = T),
            
            co2_precision = sd(co2_height, na.rm = T)/mean_co2_height*100,
            ch4_precision = sd(ch4_height, na.rm = T)/mean_ch4_height*100,
            n2o_precision = sd(n2o_height, na.rm = T)/mean_n2o_height*100) 
std_heights
```

combine std heights with unknowns

```{r}
unknown_std_heights <- gc_data_tidy %>% 
  filter(ecosystem != "STD") %>% 
  left_join(std_heights, by = "gc_run_date")
```

calculate ghg in ppm

```{r}
ghg_ppm <- unknown_std_heights %>% 
  mutate(co2_ppm = co2_height * (998/mean_co2_height),
         ch4_ppm = ch4_height * (10.2/mean_ch4_height),
         n2o_ppm = n2o_height * (1/mean_n2o_height)) 
ghg_ppm
```

write data

```{r}
write_csv(ghg_ppm, "data/tidy/230720/230720_t4_ghg_ppm.csv")
```



